\emph{Gaia} DR2 adds clues for dynamical ejection of a $\sim$$150\,M_\odot$ star

### 2018-05-24 VFTS682.tex
typo
added todo
fixed affiliations

% \title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
%                                 % ejected %% to fit in one line
% runaway star known to date}
\title{\emph{Gaia} DR2 adds clues for dynamical ejection of
  a $\sim$$150\,M_\odot$ star}
\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
  R.~P.~van~der~Marel$^{3,4}$,   I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
  V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{7,8}$,  A.~de~Koter$^{1}$,
  N.~Langer$^{9}$, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
\affil{$^{1}${\todo{check affil. num}Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
  $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
  $^{3}$ {Space Telescope Science Institute, 3700 San Martin Drive,
    Baltimore, MD 21218, USA}\\
  $^{4}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
  $^{}${Department of Physics and Astronomy, Hicks Building,
    Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
  $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada} 
  $^{}$ {School of Astronomy \& Space Science, University of the Chinese
    Academy of Sciences, Beijing 100012, China}\\
  $^{}$ {National Astronomical Observatories, Chinese Academy of
    Sciences, Beijing 100012, China}\\
  $^{}$ {Department of Physics, University of Oxford, Keble Road,
    Oxford OX1 3RH, UK} \\
  $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}   
  $^{}$ {Space Telescope Science Institute, 3700 San Martin Drive,
    Baltimore, MD 21218, USA}\\
  {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
  $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
  $^{}$ {National Research Council, Herzberg Astronomy \&
    Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
    Canada}\\
  $^{}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn, Auf dem H\"ugel 71, 53121, Bonn, Germany}
}

@@

% \todo{recheck all numbers}
\section{Introduction}
\label{sec:intro}

@@

\begin{equation}
  \label{eq:kin_age}
  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu_\mathrm{RA}^\mathrm{Gaia}} \simeq
  \frac{119.4\,\mathrm{arcsec}}{0.13\,\mathrm{mas\ yr^{-1}}} \simeq 0.9\pm\,0.6\, \mathrm{Myr} \ \ ,
\end{equation}
where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS 682 to
the core of R136 \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
As in the rest of this study, we neglect for
simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.

The kinematic age $\tau_\mathrm{kin}$ is similar or smaller than the apparent age
of the star of $1.0\pm 0.2$\,Myr from \cite{schneider:18}, which
corroborates the idea that the star is the result of a dynamical
ejection very soon in the cluster evolution. We emphasize that the
present-day surface helium abundance
\citep[$Y\simeq0.5$,][]{bestenlehner:11, rubio-diez:17} puts a lower
limit on the age corresponding to the time needed for the star to
synthesize that amount of helium, and such lower limit corresponds to
$\sim$0.9\,Myr based on the models from \cite{kohler:15}.

>>>>

% \title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
%                                 % ejected %% to fit in one line
% runaway star known to date}
\title{\emph{Gaia} DR2 adds clues for dynamical ejection of
  a $\sim$$150\,M_\odot$ star}
\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
  R.~P.~van~der~Marel$^{3,4}$,   I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
  V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{7,8}$,  A.~de~Koter$^{1}$,
  N.~Langer$^{9}$, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
\affil{$^{1}${Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
  $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
  $^{3}$ {Space Telescope Science Institute, 3700 San Martin Drive,
    Baltimore, MD 21218, USA}\\
  $^{4}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
  $^{5}${Department of Physics and Astronomy, Hicks Building,
    Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
  $^{6}$ {National Research Council, Herzberg Astronomy \&
    Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
    Canada}\\
  $^{7}$ {School of Astronomy \& Space Science, University of the Chinese
    Academy of Sciences, Beijing 100012, China}\\
  $^{8}$ {National Astronomical Observatories, Chinese Academy of
    Sciences, Beijing 100012, China}\\
  $^{9}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
    Auf dem H\"ugel 71, 53121, Bonn, Germany}\\
  $^{10}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}\\
  $^{11}$ {Department of Physics, University of Oxford, Keble Road,
    Oxford OX1 3RH, UK} \\
  $^{12}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
}

@@

\section{Introduction}
\label{sec:intro}

@@

\begin{equation}
  \label{eq:kin_age}
  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu^\mathrm{Gaia}} \simeq
  \frac{119.4\,\mathrm{arcsec}}{0.13\,\mathrm{mas\ yr^{-1}}} \simeq 0.9\pm\,0.6\, \mathrm{Myr} \ \ ,
\end{equation}
where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS 682 to
the core of R136 \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
As in the rest of this study, we neglect for
simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.

The kinematic age $\tau_\mathrm{kin}$ is similar or smaller than the apparent age
of the star of $1.0\pm 0.2$\,Myr from \cite{schneider:18}, which
corroborates the idea that the star is the result of a dynamical
ejection very soon in the cluster evolution. We emphasize that the
present-day surface helium abundance
\citep[$Y\simeq0.5$,][]{bestenlehner:11, rubio-diez:17} puts a lower
limit on the age corresponding to the time needed for the star to
synthesize that amount of helium, and such lower limit corresponds to
$\sim$0.9\,Myr based on the models from \cite{kohler:15}.
\todo{discuss position angle}
